Practical consistency management for geographically distributed MMOG servers

### 2011-05-27 textos/optpaxos/draftv2.tex
propPending

  In this case, when $now > m.ts + w(p)$, $p$ tries to initiate a consensus instance within $G$ to decide $m$. If $p$ cannot start the proposal now, $m$ is enqueued in $propPending$ for being proposed later. Before proposing $m$, $p$ checks whether $m$ -- or a message $m' : m'.ts > m.ts$ -- has already been decided in some previous consensus instance. If it has, $m$ is discarded; otherwise, $p$ starts a consensus instance proposing $m$. Once any message $m$ is decided, if $G \in m.dst$, $m$ is inserted in the $consPending$ list. Besides, $p$ sends a message $\{m, \text{`cons'}\}$ to every $p' : (\exists G' \in m.dst \setminus \{G\}) \wedge (p' \in G')$. When $\{m, \text{`cons'}\}$ is received by each $p' \in G'$, $p'$ checks whether it has ever inserted $m$ in its own $consPending$ list. If not, $p'$ inserts $m$ into $consPending$ and adjusts $barrier(G, G')$ to $m.ts$.

@@

A more formal description of the protocol is given in Algorithm \ref{algorithm:quasi}. We consider that three primitives are given: $getTime()$, which returns the current value of the local wallclock; $Propose(val, learners)$, which initiates a Paxos \cite{paxosref} instance proposing the value $val$ with the local group as the group of acceptors and $learners$ as the group of learners; and also $Learn(val)$, which is called when a Paxos instance achieves consensus. The $Learn(val)$ is called for all the processes which are supposed to learn a given proposed value once it is decided by a Paxos instance. For the sake of simplicity, we assume that, for each group, the values are learnt by consensus in the same order in which the consensus instances were initiated\footnote{This can be easily done with a sequence number given by the group's coordinator when initiating each consensus instance -- but note that we don't care if consensus instances initiated by different groups are finished in a non-deterministic order, but only those initiated in each single group; we want something like a ``FIFO learn order'', and not necessarily atomic.}.

>>>>

  In this case, when $now > m.ts + w(p)$, $p$ tries to initiate a consensus instance within $G$ to decide $m$. If $p$ cannot start the proposal now, $m$ is enqueued in $propPending$ for being proposed later. Before proposing $m$, $p$ checks whether $m$ -- or a message $m' : m'.ts > m.ts$ -- has already been decided in some previous consensus instance. If it has, $m$ is discarded; otherwise, $p$ starts a consensus instance proposing $m$. Once any message $m$ is decided, if $G \in m.dst$, $m$ is inserted in the $consPending$ list. Besides, when $m$ is decided, $p$ sends a message $\{m, \text{`cons'}\}$ to every \mbox{$p' : (\exists G' \in m.dst \setminus \{G\}) \wedge (p' \in G')$}. When $\{m, \text{`cons'}\}$ is received by each $p' \in G'$, $p'$ checks whether it has ever inserted $m$ in its own $consPending$ list. If not, $p'$ inserts $m$ into $consPending$ and adjusts $barrier(G, G')$ to $m.ts$. To ensure that, once a message $\{m, \text{`cons'}\}$ is received, there is no other message \mbox{$\{m', \text{`cons'}\}:m'.ts<m.ts$} coming from $m.src$, every $\{m, \text{`cons'}\}$ message is sent through a lossless FIFO channel\footnote{An ordinary TCP connection would be enough to provide such FIFO lossless channel.}.

@@

A more formal description of the protocol is given in Algorithm \ref{algorithm:quasi}. We consider that three primitives are given: $getTime()$, which returns the current value of the local wallclock; $Propose(I, val)$, which proposes a value $val$ for the consensus instance of id $I.id$ within the group $I.grp$; and also $Decide(I, val)$, which is called when the consensus instance $I$ finishes. $Decide(I, val)$ is called for all the processes of the group $I.grp$, when they learn that the value $val$ has been agreed upon in $I$. For the sake of simplicity, we assume that, for consensus instances within the same group, the values are decided in the same order of the instances id's\footnote{This can be easily done by delaying the callback of $Decide(I, val)$ while there is some unfinished consensus instance $I': I'.grp = I.grp \wedge I'.id < I.id$.}.
